Implementation of a Brute Force Attack on A5/1 Cryptographical Algorithm in a GPU-based Volunteering Computing Project

### 2017-02-02 Paper PaCT 2017/paper.tex
+ Section 5 complete

Among all the differrent methods of A5/1 cryptanalysis we distinguish those
which were realized in practice and allowed to reliably conduct the
cryptanalysis procedure on an non-weakened variant of algorithm. Apparently,
the first such attack was performed in 2008 with the help of the special
FPGA-based computational platform COPACOBANA\cite{COPAC_1}. In 2009 distributed
algorithms for boolean satisfiability (SAT) were used to solve several
cryptanalysis instances of stock A5/1 in a specialized grid system 'BNB-Grid'
\cite{TRUDY_ISA}. These results were further improved in 2011 \cite{SZBP}. By
the end of 2009 the 'A5/1 Cracking Project' had published rainbow tables
\cite{RAINBOW} for A5/1. Provided with 8 bursts (912 bits) of keystream these
tables allowed to find the secret key in less than a minute with more than 85\%
probabilty. Despite the huge (over 2 Tb) size, to this day these rainbow tables
provide the most practical method of A5/1 cryptanalysis.  It's main shortcoming
is that the probability of success is significantly less than 100\%. Meanwhile,
the growth of computational power of GPUs and FPGAs made practical the attack
based on reduction of search space from $2^64$ to $2^53$, which was described
by R.  Anderson in 1994. As was mentioned earlier, FPGA-based variant of the
Anderson's attack was already performed in 2008 by COPACOBANA creators. So, the
primary goal of our work is to demonstrate the viability of GPU-based variant
of the attack. Let us note that GPUs are much easier to operate than FPGAs, and
the former belong to the class of consumer-grade devices and could be found in
any modern PC, while the latter belong to the class of specialized equipment.
With the usage of the BOINC software platform \cite{BOINC_C}, this qualities of
GPUs allowed us to implement the attack in the form of a volunteering computing
project using idle computational capabilites of the project members home PCs.
Our estimates of attack's speed were based on our previous
work\cite{BUL_SEM_2016}.

@@

The described vulnerability of GSM protocol allows to build some successful
attacks on it, based on the idea of 'advanced brute force'. In particular, the
large-scale preprocessing made possible the creation of the rainbow tables,
which, provided 8 bursts of keystream, allows to determine the session key with
probability in less than a minute with $>85\%$ probability. The tables take
about 2Tb of disk space. This attack is presented in detail in \cite{RAINBOW},
and still stands among the most practical ones. Instead, we will focus on the
idea of an attack that was suggested by Ross Anderson in 1994 in a small essay
on the A5/1 cryptographic strength \cite{COPAC}. Next we describe the essence
of the Anderson's attack.

@@

Performance of the different implementations of the According to BOINC
statistics, 143 active hosts of 90 volunteers participated in the experiment.
Here by ‘active’ we mean a host which correctly processed at least one task.
Active volunteer has at least one active host. It should be noted, there is
only 1 volunteer from Russia in top-10 project volunteers list (formed
according to the amount of credits), the most of them are from United States (4
persons). In top-10 list of teams there is only one team from Russia too.

Winery~\cite{Winery} is graphical \commentontext{modeling}{modeling with one \enquote{l}, because of AE} tool.

\begin{figure}
Simple Figure
\caption{Simple Figure}
\label{fig:simple}
\end{figure}

\begin{table}
\caption{Simple Table}

\label{tab:simple}
Simple Table
\end{table}

cref Demonstration: Cref at beginning of sentence, cref in all other cases.

\Cref{fig:simple} shows a simple fact, although \cref{fig:simple} could also show something else.

\Cref{tab:simple} shows a simple fact, although \cref{tab:simple} could also show something else.

\Cref{sec:intro} shows a simple fact, although \cref{sec:intro} could also show something else.

Brackets work as designed:
<test>

The symbol for powerset is now correct: $\powerset$ and not a Weierstrass p ($\wp$).

\begin{inparaenum}
\item All these items...
\item ...appear in one line
\item This is enabled by the paralist package.
\end{inparaenum}

``something in quotes'' using plain tex or use \enquote{the enquote command}.

\section{Conclusion and Outlook}

\subsubsection*{Acknowledgments}
...

In the bibliography, use \texttt{\textbackslash textsuperscript} for ``st'', ``nd'', ...:
E.g., \enquote{The 2\textsuperscript{nd} conference on examples}.
When you use \href{https://www.jabref.org}{JabRef}, you can use the clean up command to achieve that.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs03}
\bibliography{paper}

All links were last followed on October 5, 2014.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Among all the differrent methods of A5/1 cryptanalysis we distinguish those
which were realized in practice and allowed to reliably conduct the
cryptanalysis procedure on an non-weakened variant of algorithm. Apparently,
the first such attack was performed in 2008 with the help of the special
FPGA-based computational platform COPACOBANA\cite{COPAC_1}. In 2009 distributed
algorithms for boolean satisfiability (SAT) were used to solve several
cryptanalysis instances of stock A5/1 in a specialized grid system 'BNB-Grid'
\cite{TRUDY_ISA}. These results were further improved in 2011 \cite{SZBP}. By
the end of 2009 the 'A5/1 Cracking Project' had published rainbow tables
\cite{RAINBOW} for A5/1. Provided with 8 bursts (912 bits) of keystream these
tables allowed to find the secret key in less than a minute with more than 85\%
probabilty. Despite the huge (over 2 Tb) size, to this day these rainbow tables
provide the most practical method of A5/1 cryptanalysis.  It's main shortcoming
is that the probability of success is significantly less than 100\%. Meanwhile,
the growth of computational power of GPUs and FPGAs made practical the attack
based on reduction of search space from $2^64$ to $2^53$, which was described
by R.  Anderson in 1994 \cite{ANDERSON}. As was mentioned earlier, FPGA-based
variant of the Anderson's attack was already performed in 2008 by COPACOBANA
creators.  So, the primary goal of our work is to demonstrate the viability of
GPU-based variant of the attack. Let us note that GPUs are much easier to
operate than FPGAs, and the former belong to the class of consumer-grade
devices and could be found in any modern PC, while the latter belong to the
class of specialized equipment.  With the usage of the BOINC software platform
\cite{BOINC_C}, this qualities of GPUs allowed us to implement the attack in
the form of a volunteering computing project using idle computational
capabilites of the project members home PCs.  Our estimates of attack's speed
were based on our previous work\cite{BUL_SEM_2016}.

@@

The described vulnerability of GSM protocol allows to build some successful
attacks on it, based on the idea of 'advanced brute force'. In particular, the
large-scale preprocessing made possible the creation of the rainbow tables,
which, provided 8 bursts of keystream, allows to determine the session key with
probability in less than a minute with $>85\%$ probability. The tables take
about 2Tb of disk space. This attack is presented in detail in \cite{RAINBOW},
and still stands among the most practical ones. Instead, we will focus on the
idea of an attack that was suggested by Ross Anderson in 1994 in a small essay
on the A5/1 cryptographic strength \cite{ANDERSON}. Next we describe the
essence of the Anderson's attack.

@@

According to BOINC statistics, 143 active hosts of 90 volunteers participated
in the experiment.  Here by ‘active’ we mean a host which correctly processed
at least one task.  Active volunteer has at least one active host. It should be
noted, there is only 1 volunteer from Russia in top-10 project volunteers list
(formed according to the amount of credits), the most of them are from United
States (4 persons). In top-10 list of teams there is only one team from Russia
too.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related works.}\label{sec:related}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As was noted before, at the time of writing of our article the A5/1 algorithm
remains one of the most popular research subjects among cryptography
specialists, along with DES and RSA algorithms. The exact authorship of this
algorithm is unknown, but some experts argue that chances are high that it's
creation should be attributed to a French special agencies. Numerous sources
describe the events that led to it's structure leaked to general public. In
fact, it's internal workings were already known to research community in 1994:
R.Anderson's note (a mailing list message) written in that period was one of
the firs works on A5/1 cryptanalysis. The complete knowledge of the A5/1 were
acquired in 1999 as a result of a reverse engineering of a mobile phone.

Earlier we mentioned Anderson's attack (described by Ron Anderson in a short
note in 1994 \cite{ANDERSON}) was the first attack on A5/1 with compexity less
than that of a trivial brute-force walk over the whole keyspace. Shortly
afterwards J.Golic suggested the "Alleged" attack on the A5/1, based on a
linearization of the equations describing A5/1 \cite{GOLIC}. The Golic's attack
estimated complexity is $C\dot2^{40}$, where $C$ denoting the complexity of
solving the system of linear equations over $GF(2)$ with a very substanial
number of dimensions. The next widely known attack on the A5/1 was presented in
\cite{BSW}. To speed up the A5/1 algorithm it's authors used the technique of
precomputation of LFSRs we mentioned in \cref{sec:bitsl}. It's worth to note
that the attack presented in \cite{BSW}, along with those presented in
\cite{BIKHAM2000}, \cite{EKDAHL2003}, \cite{BARKAN2008} \todo{Check the actual
articles! Not sure if these point to the right ones.}, require substanial
lengths of known keystream (several seconds at best). These attacks can be
deemed realistic, but they do not demonstrate vulnerability of the GSM security
protocols as evidently, as those belonging to the "advanced brute force" class,
that we will discuss below. For the rest of the section we consider an attack
practical if it's real-life implementation is able to persistently find the
secret key by analyzing no more than 2 frames (912 bits) of the keystream, the
amount that could always be extracted from a call in a GSM network by
exploiting the protocol vulnerability described in \Cref{sec:alg}.

It seems that the first practical attack ('practical' in a sense we established
above) was presented in \cite{COPAC_1}. It's authors implemented the optimezed
variant of the Anderson's attack on specialized computational device of their
own design, assembled from 120 "Xilinx Spartan 3" FPGAs. The authors state
\cite{COPAC_1} that the attack took about 6 hours to complete. It's worth
mentioning that the COPACOBANA architecture was used for cryptanalytical
research of several other ciphers, e.g. DES \cite{COPAC_2}.

The first estimates of the time required for the cryptanalysis of A5/1 on a
computational cluster in the form of a boolean satisfiability problem (SAT)
\cite{HANDBOOK} were presented in \cite{PACO-2008}. SAT-based cryptanalysis is
considered a perspective direction of cryptanalysis that seems to be on the
rise for the last decade. It operates within a paradigm that the
cryptanalytical problem could be effeciently represented in the form of a SAT
problem. This could be achieved with the help of special translators such as
\cite{CRYPTOL}, \cite{URSA}, \cite{TRANSALG}. After the problem was translated
to a SAT form, it can be efficiently solved by modern SAT algorithms that could
run in a distributed computing environment. The cryptanalysis time of A5/1
estimated in \cite{PACO-2008} was realistic, but since the exclusive usage of
the whole computational cluster for prolonged period of time was not an option,
the actual attack was performed a year later in a specialized grid-system
\cite{TRUDY_ISA}. In the following years these results were improved - the
search space of $2^{31}$ SAT instances was processed completely and collisions
(different secret keys generating the same keystream) have been found
\cite{SZBP}. At the same period it became apparent that the organizational
requirements of this kind of cryptographical attack fits well into the ideology
of volunteering computing projects \todo{cite what?}. To bring this line of
thought to life in 2012 SAT@home \cite{SATHOME} volunteering computing project
was incepted. Through 4 years of it's activity a number of hard combinatorial
problems were solved within the SAT paradigm. Among those were several dozens
of cryptanalysis of A5/1 generator problems (only one burst - 114 bits of
keystream - was used each time). These results were published in
\cite{PACT_2015}, \cite{SPRINGER}.

The work \cite{COPAC_2} included the estimates of creation time and disk space
that would be required to create the rainbow-tables that would make possible to
'break' the A5/1 on an average PC in several minutes. \cite{COPAC_2} estimated
these to occupy about 7 Tb of disk space. "The A51 Cracking Project" put the
significantly more compact (2 Tb) tables into public domain at the end of 2009
\todo {cite what?}. By analysing 2 frames (912 bits) of known keystream with
the help of this tables one can restore the secret key with probability of
success over 85\%. This method of cryptanalysis of the A5/1 algorithm could be
assumed to be the most practical. The fact that it does not provide 100\%
guarantee of success from analysis of 2 frames does not matter in real
circuimstances, where much longer samples of keystream are available to the
attacker. It should be noted that if the tables are lost, it will require a
vast amount of computational resources to generate them again.

Finally, the attack we presented in this paper does not lose on the practical
side to the rainbow-tables attack, because it allows to complete the
cryptanalysis with 100\% success rate in a period of several hours in a small
volunteering computation project. Moreover, the attack requires only 64 bits of
known keystream. Volunteering computing projects are rising in popularity and
publicity in the last years, so any "hot" problem (like a cryptanalysis of one
of the most commonly used ciphers) immediately gets volunteers attention.

As a conclusion, we would like to express the hope that the efficiency of our
implementation of attack on A5/1 generator will serve as yet another strong
argument against the usage of this algorithm in the transfer of sensible data
through GSM networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs03}
\bibliography{paper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
